---
title: "Let's practice 11"
date: "`r Sys.Date()`"
output:
 html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    toc_float:
      collapsed: true
    code_folding: hide
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, include=TRUE, warning = FALSE, message=FALSE)
```

# Flow cytometry data analysis

We perform analysis of flow cytometry data from the flowRepository accession FR_FCM_Z3WR. Data was analyzed with `r sessionInfo()$R.version$version.string`. Most analysis steps were performed using the following packages: flowCore (v`r packageVersion("flowCore")`), CATALYST (v`r packageVersion("CATALYST")`). 

```{r libraries}

# load libraries
library(flowCore)
library(CATALYST)
library(cowplot)

```

## A clean flowSet output by flow AI is imported. 

```{r import-flowSet-and-panel}

# 1) load the flowSet object from exercise 3
load("../course_datasets/FR_FCM_Z3WR/fcs_clean.RData")

# 2) Downsample to 2'000 cells 

# source downsampling function
source("../course_datasets/function_for_downsampling_flowSets.R")

# downsample
set.seed(1234)
fcs_small <- Downsampling_flowSet(fcs_clean,samplesize = 2000)

# 3) Create a sce object from the downsampled flowSet
# We need the panel data frame
panel <- read.csv("../course_datasets/FR_FCM_Z3WR/panel_with_marker_classes.csv")

colnames(panel)<-c("dd")
# We also need a metadata dataframe
md <- pData(fcs_small)

# create sce
sce <- prepData(fcs_small, 
                md = md,
                md_cols = list(file="name", id = "name", factors = "time_point"),
                panel = panel,
                panel_cols = list(channel = "channels", antigen = "antigen", class = "marker_class"),
                transform = FALSE,
                FACS=TRUE,
                features = panel$channels[panel$marker_class!="none"])

# change the assay name to "exprs" 
assayNames(sce) <- "exprs"

```

## The UMAP is computed using default parameters 

Parameters = n_neighbors=15 and min_dist = 0.01.

```{r umap}
# 4) UMAP with default parameters (n_neighbors=15 and min_dist = 0.01)
set.seed(1601)
sce <- runDR(sce, 
             assay = "exprs", 
             dr = "UMAP", 
             features = "type", 
             cells = NULL)


```

## UMAP of down-sampled cells

### CD3 expression 
```{r umap-1, fig.align='center', fig.width=10, fig.height=10}

# plot
plotDR(sce,
       dr =  "UMAP",
       assay = "exprs", 
       color_by="CD3", 
       facet_by = "time_point" )

```

## UMAP of down-sampled cells

### CD8 and CD4 expression

Using plot_grid()

```{r umap-2, fig.align='center'}

# plot CD4
p1<-plotDR(sce,
       dr =  "UMAP",
       assay = "exprs", 
       color_by="CD4")

# plot CD4
p2<-plotDR(sce,
       dr =  "UMAP",
       assay = "exprs", 
       color_by="CD8")

plot_grid(p1, p2, nrow = 2)

```

## UMAP of down-sampled cells

### CD19 and CD56 expression {.unnumbered}

Providing a vector with several antigens to the plotDR() function, this also creates a grid.

```{r umap-3, fig.align='center'}

# provide 2 genes
plotDR(sce,
       dr =  "UMAP",
       assay = "exprs", 
       color_by=c("CD19", "CD56"))


```

<!-- This section is an R Markdown comment

here what eve

You can comment out full parts and ignore them completely from your final knitted document

```{r cars, include=TRUE, eval=TRUE, echo=TRUE}
summary(cars)
```

-->

# Information about R and package versions
```{r, session-info, results='markup', warning=TRUE, message=TRUE}
sessionInfo()
```
